\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Warsaw}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}

% Avoid compiling twice
\usepackage{bookmark}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\newcommand{\y}{\ensuremath{\mathbbm{y}}}

\title[k -- sparsity]
{k -- sparsity}

\subtitle
{Group Testing}

\author[Amit, Vikraman]
{Amit Kumar Sinhababu\inst{1} \and Vikraman Choudhury\inst{1}}

\institute[Indian Institute of Technology Kanpur]
{
  \inst{1}
  Department of Computer Science and Engineering\\
  Indian Institute of Technology Kanpur
}

\date[CS719]
{Data Streaming Algorithms and Systems}

\subject{Theoretical Computer Science}

\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}


\section{Motivation}

\begin{frame}{Coding Theory}
  Coding Theory in Data Streaming
\end{frame}

\section{Group Testing}

\begin{frame}{Applications of Group Testing}
  \begin{itemize}
  \item Computational Biology
  \item DNA Library screening
  \item Multiple access control protocols
  \item Data pattern mining
  \item Data Forensics
  \end{itemize}
\end{frame}

\section{Connections to other problems}

\subsection{Compressed Sensing}

\subsection{Bipartite Expander Graphs}
\begin{frame}
  A $W$-left regular bipartite graph $[N]x[W] \rightarrow [T]$ is an
  $(N, W, T, D, (1-\epsilon)W)$ expander if every subset $S \subset [N]$
  of size at most $D$ has a neighborhood (denoted by $\Gamma(S)$) of size
  at least $(1-\epsilon)|S|W$.
\end{frame}

\begin{frame}
  Let $G$ be a $(n,w,t,2,w(1-\epsilon))$-expander. Then $M_G$ is a $d$-disjunct matrix.
\end{frame}

\begin{frame}
  Let $G$ be a $(n,w,t,2d,w/2+1)$-expander.
  Then $M_G$ is a $(d,d)$-list disjunct matrix.
\end{frame}

\subsection{Bipartite Disperser Graphs}
\begin{frame}
  A $D$-left regular bipartite graph $[N]x[D] \rightarrow [T]$ is an
  $(N, L, T, D, \epsilon)$ disperser if every subset $S \subset [N]$
  of size at least $L$ has a neighborhood of size at least $\epsilon|T|$.
\end{frame}

\begin{frame}
  Let $G$ be a $(n,l,t,\epsilon t/(d+1),\epsilon)$-disperser.
  Then $M_G$ is a $(d,l)$-list disjunct matrix.
\end{frame}

\subsection{Sparsity Separator}
\begin{frame}
  Given a stream of $m$ items from the domain $[n]$, let $f$ be the vector of
  frequencies, i.e., for every $i \in [n]$, $f_i$ denotes the number of occurences
  of $i$ in the stream.

  A $(d,l)$-sparsity separator is a data structure which can determine if the
  frequency vector corresponding to a stream has at most $d$ non-zero entries
  or it has at least $l$ non-zero entries.
\end{frame}

\begin{frame}
  Any $(d,l-d)$-list disjunct $txn$ matrix $M$ can be used to build a
  $(d,l)$-sparsity separator structure.
\end{frame}

\subsection{Heavy Hitter}

\subsection{Tolerant Testing}
\begin{frame}
  Given a received word $\y \in \Sigma^n$, we need to decide
  if \y\ is at a distance at most $e_1$ from some codeword or
  if it has distance at least $e_2 > e_1$ from every codeword in $C$.
\end{frame}

\end{document}
